Gastrointestinal I

### 2017-09-19 git1.rmd
mc

---
title: "Module II"
author: "Aaron Khoo"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  pdf_document
documentclass: article
classoption: a4paper
fontsize: 11pt
urlcolor: blue
---

@@

-  CVS: ruptured AAA, MI
-  GIT: oesophagus, stomach, bile ducts, liver, large intestine, pancreas, spleen, small intestine, *peritoneum*
-  Neuro: radicular pain
-  MSK: rectus abd
-  Resp: 
-  Renal: nephrolithiasis, pyelo
-  Repro: testicular torsion
-  Endocrine:

@@

1.  Abdo/chest pain
2.  Altered ingesting of food (from nausea, vomiting, dysphagia, odynophagia [painful swallowing], anorexia)
3.  ALtered bowel movements (diarrhoea or constipation)
4.  GI bleeding

@@

-  Somatic = structures supplied by spinal nerves, includes the parietal pleura and peritoneum. Thus perforated duodenal ulcer and generalised peritonitis can cause diffuse abdo pain (T6-T12). 
-  Visceral: sympathetic component of the ANS (rami communicans => sympathetic chain => sphlancnic nerves [preganglionic efferent stimulatory signals; greater 59, lesser 1011, least 12, lumbar 12] => prevertebral vascular ganglia and plexues => postganglionic efferent fibres follow individual arteries to viscus

@@

@@

@@

Normal: balance between gastric acid secretion and gastroduodenal defense (mucus and bicarbonate secretion, ion pumps to remove excess H+ *intracellularly*, tight junctions, cellular restitution, epithelial renewal), restitution where healthy cells migrate to epithelium. Patho: agressive factors such as NSAIDs or *H. pylori*, alcohol, bile salts, acid, pepsin can alter this balance by allowing back diffusion of H+ => epithelial damage. Smoking convers increased risk of gastric ulcer

@@

1.  Cephalic phase: sight, smell, taste, thoughts => medulla => vagal ouput => submucosal plexus => increases secretion from gastric glands and stomach motility
2.  Gastric phase: food reaching stomach => stretch and chemoreceptors => secretion of gastrin by G cells
3.  Intestinal phase: stimulation of intestinal receptors => secretion of duodenal hormones: *secretin* reducers gastric secretion, cholecystokinin inhibits gastric emptying

NSAIDs: Non-selective can damage the gastric and duodenal mucosal barrier and are associated with an increased risk of upper gastrointestinal ulceration, bleeding and perforation.

@@

-  Intro
-  General inspection: mental state (withdrawal symptoms), colour (jaundice, pallour, bronzed skin), hydration
-  Inspect hands and upper limb:clubbing, nail changes (iron deficiency, chronic liver disease), hepatic flap (asterixis from hepatic encephalopathy or CO2 retention or renal failure)
-  Radial pulse, RR, BP
-  Face inspection: eyes, glands, mouth
-  Neck lymph nodes
-  Inspection chest and abdomen: chest, abdomen = {masses/distension, stretch marks, colour changes, prominent veins, pulsations, scars}
-  Palpate the abdomen: 9 locations, look for tenderness at McBurneys point (acute appendicitis), Rovsing's sign is rebound tenderness (localised peritonism), Murphy's sign is tenderness under right costal margin during inspiration (acute cholecystitis), Courvoisier's law is palpable non-tender gallbladder under right costal marging + pt jaundiced (gallstones unlikely so more likely cancer of head of pancreas)
-  Liver Ex: palpate, percuss
-  Spleen: palpate
-  Kidney: palpate
-  Abdo aorta
-  Ascites
-  Auscultate bowel signs: active gurgling signs called borborygmi (diarrhoeal states), shake pt and here succussion splash which is liquid+air within luminal organ (normal if eaten, path if pt is nil bu mouth for >6 hours and clear fluids >2 hours), auscultate for bruits
-  Lower limbs: inspection and palpation

@@

@@

-  Bloods: FBC, CHEM20, amylase, lipase
-  ECG?
-  *H. pylori tests*: 
-  Digital rectal exam
-  Endoscopy: preferred as ulcers may be malignant, also follow up to ensure healing

@@

Peptic ulcers are defects in the gastric or duodenal mucosa that extend through the muscularis mucosa. Mucosal injury (+ PUD) occurs when balance is disrupted. Factors such as NSAIDs, *H. pylori*, alcohol, bile salts, acid, pepsin can alter the mucosal defence allowing *back diffusion* of H+ => epitheliel injury. *Acute* = no signs of fibrosis vs chronic, erosions do *not* penetrate the msucularis mucosae.

@@

-  All pt with proven ulcers and H. pylori positie have eradication as primary therapy. PPI taken simultaneously with two antibiotics (from amoxicillin, clarithromycin and metronidazole) for 7 days. Success in 80-90% of people, depends on antibiotic resistance, compliance due to side-effects (diarrhoea common, flushing and vomiting, nausea, abdo cramps, headache, rash).
-  General: stop cigarettes, aspiring, and NSAIDs
-  Surgery if indicated: persistent ulceration, recurrent ulceration, gastric outflow obstruction, perforation/haemorrhage

@@

-  PPI
    -  Esomeprazole: blocks H+ secretion so stomach pH is increased, reducing irritation of stomach epithelium, allowing healing
-  Antibiotics
    -  Amoxycillin: 3rd generation aminopenicillin, $\beta$-lactam which inhibits cell wall synthesis. 
    -  Clarithromycin: macrolide so it is a 50s inhibitor, inhibiting bacterial protein synthesis

>>>>

---
title: "Gastrointestinal I"
author: "Aaron Khoo"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  pdf_document
documentclass: article
classoption: a4paper
fontsize: 11pt
urlcolor: blue
---

@@

-  CVS: ruptured AAA, MI
-  GIT: oesophagus, stomach, bile ducts, liver, large intestine, pancreas, spleen, small intestine, *peritoneum*
-  Neuro: radicular pain
-  MSK: rectus abd
-  Resp: pneumonia can sometimes present
-  Renal: nephrolithiasis, pyelo
-  Repro: testicular torsion
-  Endocrine:

@@

1.  Abdo/chest pain
2.  Altered ingesting of food (from nausea, vomiting, dysphagia, odynophagia [painful swallowing], anorexia)
3.  Altered bowel movements (diarrhoea or constipation)
4.  GI bleeding

@@

-  Somatic: structures supplied by spinal nerves, includes the parietal pleura and peritoneum. Thus perforated duodenal ulcer and generalised peritonitis can cause diffuse abdo pain (T6-T12). 
-  Visceral: sympathetic component of the ANS (rami communicans => sympathetic chain => sphlancnic nerves [preganglionic efferent stimulatory signals; greater 59, lesser 1011, least 12, lumbar 12] => prevertebral vascular ganglia and plexues => postganglionic efferent fibres follow individual arteries to viscus 
-  Example of visceral => somatic: appendicitis starts off as general then localises as it is visceral then starts affecting peritoneum (somatic pain)

@@

*Look at patient, not translator during Hx taking*

@@

Pancreatitis: usually caused by heavy drinking

Black foul smelling stool: malaena from blood

@@

Normal: balance between gastric acid secretion and gastroduodenal defense (mucus and bicarbonate secretion, ion pumps to remove excess H+ *intracellularly*, tight junctions, cellular restitution, epithelial renewal), restitution where healthy cells migrate to epithelium. Patho: agressive factors such as NSAIDs or *H. pylori*, alcohol, bile salts, acid, pepsin can alter this balance by allowing back diffusion of H+ => epithelial damage. Smoking confers increased risk of gastric ulcer

@@

1.  Cephalic phase: sight, smell, taste, thoughts => medulla => vagal ouput => submucosal plexus => increases secretion from gastric glands and stomach motility
2.  Gastric phase: food reaching stomach => stretch and chemoreceptors => secretion of gastrin by G cells
3.  Intestinal phase: stimulation of intestinal receptors => secretion of duodenal hormones: *secretin* reduces gastric secretion, cholecystokinin inhibits gastric emptying

NSAIDs: Non-selective can damage the gastric and duodenal mucosal barrier and are associated with an increased risk of upper gastrointestinal ulceration, bleeding and perforation. Prostaglandins => decrease gastric acid

Melaena: black malodourous stool is suggestive of upper GI bleeding

@@

-  Intro
-  General inspection: mental state (withdrawal symptoms), colour (jaundice, pallour, bronzed skin), hydration
-  Inspect hands and upper limb:clubbing, nail changes (iron deficiency, chronic liver disease), hepatic flap (asterixis from hepatic encephalopathy or CO2 retention or renal failure)
-  Radial pulse, RR, BP
-  Face inspection: eyes, glands, mouth
-  Neck lymph nodes
-  Inspection chest and abdomen: chest, abdomen = {masses/distension, stretch marks, colour changes, prominent veins, pulsations, scars}
-  Palpate the abdomen: 9 locations, look for tenderness at McBurneys point (acute appendicitis), Rovsing's sign is rebound tenderness (localised peritonism), Murphy's sign is tenderness under right costal margin during inspiration (acute cholecystitis), Courvoisier's law is palpable non-tender gallbladder under right costal marging + pt jaundiced (gallstones unlikely so more likely cancer of head of pancreas)
    -  Use stethoscope if you wanna try get rid of guarding
-  Liver Ex: palpate, percuss
-  Spleen: palpate
-  Kidney: palpate
-  Abdo aorta
-  Ascites
-  Auscultate bowel signs: active gurgling signs called borborygmi (diarrhoeal states), shake pt and here succussion splash which is liquid+air within luminal organ (normal if eaten, path if pt is nil bu mouth for >6 hours and clear fluids >2 hours), auscultate for bruits
-  Lower limbs: inspection and palpation

@@

He doesn't look cachexic (malignancy, malabsorptive)

@@

-  Bloods: FBC, CHEM20, amylase, lipase
-  ECG?
-  *H. pylori tests*: 
-  Digital rectal exam
-  Endoscopy: preferred as ulcers may be malignant, also follow up to ensure healing
-  USS: more likely if he has cholecystitis or biliary colic

@@

Peptic ulcers are defects in the gastric or duodenal mucosa that extend through the muscularis mucosa. Mucosal injury (+ PUD) occurs when balance is disrupted. Factors such as NSAIDs, *H. pylori*, alcohol, bile salts, acid, pepsin can alter the mucosal defence allowing *back diffusion* of H+ => epithelial injury. *Acute* = no signs of fibrosis vs chronic, erosions do *not* penetrate the muscularis mucosae.

@@

-  All pt with proven ulcers and H. pylori positive have eradication as primary therapy. PPI taken simultaneously with two antibiotics (from amoxicillin, clarithromycin and metronidazole) for 7 days. Success in 80-90% of people, depends on antibiotic resistance, compliance due to side-effects (diarrhoea common, flushing and vomiting, nausea, abdo cramps, headache, rash).
-  General: stop cigarettes, aspirin, and NSAIDs
-  Surgery if indicated: persistent ulceration, recurrent ulceration, gastric outflow obstruction, perforation/haemorrhage

@@

-  PPI
    -  Esomeprazole: blocks H+ secretion so stomach pH is increased, reducing irritation of stomach epithelium, allowing healing. Irreversible blockage of H/K ATP-ase
-  Antibiotics
    -  Amoxycillin: 3rd generation aminopenicillin, $\beta$-lactam which inhibits cell wall synthesis. 
    -  Clarithromycin: macrolide so it is a 50s inhibitor, inhibiting bacterial protein synthesis
